---
title: "Chemotaxis - multiple strains"
output:
word_document: default
html_document:
df_print: paged
pdf_document: default
---

## Independent Investigations of C. elegant Chemotaxis	

Congratulations on completing your proposal!  Please proceed to data analysis only after approval of your proposal from your instructor.

We have carried out the chemotaxis assays to 1% isoamyl alcohol for all of the mutants.  The data for each of the strains are loaded as individual .csv files with the tutorial.  Follow the directions to load the files for the strains you are proposing to analyze.  You will then use a command to combine each of these datafiles into one data frame.  Then you will proceed with the code for further calculations as you have done previously.


First, load the tidyverse:
```{r}
library(tidyverse)

```



```{r}


# To load the datasets for each strain you are analyzing, adapt the following code for the appropriate data set.  You will need a line of code for each dataset you need to load.  

#wt<-read_csv(“wt.csv");


#Then print the each dataset to make sure it loaded properly.
#wt

#To combine the datasets into one data frame, adapt the following code. Be sure to list all of the datasets you are comparing in the parentheses.

#combined<-rbind(wt, )

#Then print the combined dataset to make sure that the data for all the strains you are interested in are combined.
#combined
```



Like last couple of labs, we want to determine the Chemotaxis Index. 
The formula is:

ChemI= ((T1+T2)-(C1+C2))/(T1+T2+C1+C2)

We can add a column to our data frame by using the "mutate" function. "Mutate" allows you to create a new column of values based upon data already in the data frame. So using our example above:

```{r}


#df<-mutate(.data=combined, ChemI=((T1+T2)-(C1+C2))/(T1+T2+C1+C2))

#print out our updated df dataframe
#df

```

Now that we have the Chemotaxis Index for each plate, we can use group by/summarise to calculate the average, standard deviation, n, and upper and lower confidence intervals:

### Calculate means for the different strains.

```{r}
#meanChemI<-df %>% group_by(strain) %>% summarise(avg=mean(ChemI), StandardDeviation=sd(ChemI), n=n(), upperCI=avg+((qnorm(0.975)*StandardDeviation)/sqrt(n)), lowerCI=avg-((qnorm(0.975)*StandardDeviation)/sqrt(n)))

#meanChemI


```

### Now make a barplot using ggplot
```{r}
#ggplot(data=meanChemI, mapping=aes(x=strain, y=avg))+geom_bar(stat="identity")


```
# Take a look at your plot.  Is this the order you want the data presented?  The code has plotted the data in the order it appeared in the meanChemI data table, daf-19 followed by wt.  Generally, the data for the wild-type strain should appear first on the x-axis followed by any mutant strains. To make this change, use the following code to determine the order of the x-axis and then make the plot again.

```{r}
#meanChemI$strain<-factor(meanChemI$strain, levels=c("wt", "daf-19"))

#ggplot(data=meanChemI, mapping=aes(x=strain, y=avg))+geom_bar(stat="identity")



```


# Improve the bargraph with labels for the axes.  You do not need a title for your graph in R.  For the chemotaxis project, your final format for presenting your data will be an oral presntation with a slideshow. You will add a title later when you generate Slides to present your results.

#The theme_minimal code is a choice for how the graph background and axes look. If you want to play with how the graph looks, you can do a websearch for geom_bar themes and locate other code options to designate a different presentation of the data.
```{r}
#ggplot(data=meanChemI, mapping=aes(x=strain, y=avg))+geom_bar(stat="identity")+labs(x="your x-axis label", y="your y-axis label", fill = "Strain") + theme_minimal()+ylim(0,1)

```


### Plot a bar graph with 95% Confidence interval used for error bars and adjust the y-axis limits to extend from -1 to +1.

```{r}
#ChemIPlot <- ggplot(data=meanChemI, aes(strain, avg)) +
#geom_col() +
#geom_errorbar(aes(ymin = lowerCI, ymax = upperCI), width=0.2)

#ChemIPlot + labs(x="your x-axis label", y="your y-axis label", fill = "Strain") + theme_minimal() + ylim(-1,1)


```

### Save the plot
```{r}
ggsave("plot.png", height=5, width=8)


```


Now that you have plotted your data, it is time to do some statistical calculations to see if the mean chemotaxis index for each strain is significantly different than the others.  Previously, with only two strains, this was done with a Welch’s t-test.  For comparing multiple strains, we will use a one-way anova (analysis of variance) across all data looking at the means and the variance of the data to determine the p-value for the probability that these data are not significantly different from each other.  A p-value less than 0.05 suggests that the differences are due to real differences between the strains in the assay rather than random chance. It's worth noting that as more comparisons are made, the probability of a "false-positive" increases

For a detailed explanation of anova, visit this website. (http://www.biostathandbook.com/onewayanova.html).

```{r}

#Run the following code to run the anova analysis of your dataset.

#res.aov<-aov(ChemI ~ strain, data=df)

#Then the following line to see the results of the analysis.
#summary(res.aov)


#This gives you a p-value for your entire dataset. If you reject the null hypothesis that all the means are equal (p<0.05), you'll probably want to look at the data in more detail. One common way to do this is to compare different pairs of means and see which are significantly different from each other.  For this we will use the Tukey–Kramer method with which the minimum significant difference is calculated for each pair of means. It depends on the sample size in each group, the average variation within the groups, and the total number of groups. 
```

Run the following code to calculate a p-value for each pairwise comparison.

```{r}
#TukeyHSD(res.aov)

```


